``Interestingness'' in Mathematical Theory Exploration

### 2019-10-16 PhDThesis/clustering.tex
TODOs

% TODO: Say what's in this chapter
In this chapter we consider how the contents of a given signature can be
analysed, such that subsequent exploration can be performed more efficiently. We
frame this analysis as a task of \emph{unsupervised machine learning}, and
describe how well-known \emph{clustering algorithms} can utilise the syntax of
definitions to determine their similarity. This requires a detailed description
of Haskell definition syntax, in particular of the ``Core'' intermediate
language of the GHC compiler. We identify the difficulty posed by references,
and how these can be made less opaque through the use of \emph{recurrent
  clustering}.

@@

To improve the performance of theory exploration tools, we need to avoid
brute-force search in favour of a more informed strategy.
\iffalse
TODO: Alison: This makes it sound as though current TE tools all use brute force
(whereas all the tools I know already use informed strategies).
MY THOUGHTS: ``more informed'' is a bit of weasel-wording from me; implying that
they're not informed and that we're making them informed; whereas they're
actually already somewhat informed, and we're trying to add to that. I agree
that this weasel-wording should be replaced!
\fi
This, in turn,
requires more information \iffalse TODO: Alison: More than what? \fi about the
terms in the given signature. One
generally available source of information is the syntax of a term's definition.

@@

  \caption{Simplified syntax of GHC Core in BNF style. $[]$ and $(,)$ denote
    repetition and grouping, respectively \iffalse TODO: Alison: Include examples? \fi. Other forms of literal, like machine
    words, are treated in a similar way and have been omitted for brevity.}
  \label{fig:coresyntax}
\end{figure}

@@

\begin{figure}
  \begin{subfigure}[haskell]{\textwidth}
    \underline{\texttt{Haskell definitions}}
    \iffalse
    TODO: Alison: Big space
    MY THOUGHTS: I wanted to align things more neatly...
    \fi
    \begin{haskell}
plus    Z  y = y
plus (S x) y = S      (plus x y)

>>>>

\iffalse TODO: Alison: Each chapter should have a description at the start (see
bucketing ``in this chapter'') \fi

In this chapter we consider how the contents of a given signature can be
analysed, such that subsequent exploration can be performed more efficiently. We
frame this analysis as a task of \emph{unsupervised machine learning}, and
describe how well-known \emph{clustering algorithms} can utilise the syntax of
definitions to determine their similarity. This requires a detailed description
of Haskell definition syntax, in particular of the ``Core'' intermediate
language of the GHC compiler. We identify the difficulty posed by references,
and how these can be made less opaque through the use of \emph{recurrent
  clustering}.

@@

To improve the performance of theory exploration tools, we need to provide their
search strategies with more than just type information about the terms in the
given signature, rather than treating them as ``black boxes''. One generally
available source of information is the syntax of a term's definition.

@@

  \caption{Simplified syntax of GHC Core in BNF style. $[]$ denotes repetition
    (zero or more occurrences). Other forms of literal, like machine words, are
    treated in a similar way and have been omitted for brevity.}
  \label{fig:coresyntax}
\end{figure}

@@

\begin{figure}
  \begin{subfigure}[haskell]{\textwidth}
    \underline{\texttt{Haskell definitions}}
    \begin{haskell}
plus    Z  y = y
plus (S x) y = S (plus x y)
